
---
title: "`r params$report_title`"
author: "AGB Microbiome Analysis Platform"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
  pdf_document:
    toc: true
    number_sections: true
    latex_engine: xelatex
  word_document:
    toc: true
    reference_docx: NULL
params:
  report_title: "Microbiome Analysis Report"
  report_comments: ""
  selected_sections: ["summary", "alpha", "beta"]
  sample_summary: NULL
  taxonomy_summary: NULL
  alpha_stats: NULL
  plot_files: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE, 
  warning = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 300,
  out.width = "100%"
)

# Load required libraries
suppressPackageStartupMessages({
  library(ggplot2)
  library(knitr)
  library(dplyr)
  library(kableExtra)
})

# Helper function to check if file exists and is not empty
file_exists_and_valid <- function(filepath) {
  if (is.null(filepath) || filepath == "" || is.na(filepath)) {
    return(FALSE)
  }
  file.exists(filepath) && file.info(filepath)$size > 0
}

# Helper function to format numbers
format_number <- function(x, digits = 2) {
  if(is.numeric(x)) {
    formatC(x, format = "f", digits = digits, big.mark = ",")
  } else {
    as.character(x)
  }
}
```

---

**Report Generated:** `r format(Sys.time(), "%B %d, %Y at %I:%M %p")`

```{r comments, results='asis'}
if (!is.null(params$report_comments) && params$report_comments != "") {
  cat("\n**Analysis Comments:** ", params$report_comments, "\n\n")
}
```

---

# Executive Summary

This comprehensive microbiome analysis report presents findings from high-throughput sequencing data analysis. The report includes diversity metrics, taxonomic composition, and statistical comparisons between sample groups.

```{r summary-stats, results='asis'}
if(!is.null(params$sample_summary)) {
  cat("## Dataset Overview\n\n")
  
  sample_info <- params$sample_summary
  
  cat("- **Total Samples Analyzed:** ", sample_info$total_samples, "\n")
  cat("- **Control Samples:** ", sample_info$total_controls, "\n")
  cat("- **Collection Period:** ", sample_info$collection_period, "\n")
  cat("- **Age Range:** ", sample_info$age_range, " years\n")
  
  if(!identical(sample_info$gender_dist, "N/A")) {
    cat("- **Gender Distribution:** ")
    gender_text <- paste(names(sample_info$gender_dist), 
                        sample_info$gender_dist, 
                        sep = ": ", collapse = ", ")
    cat(gender_text, "\n")
  }
  
  cat("\n")
}

if(!is.null(params$taxonomy_summary)) {
  cat("## Taxonomic Overview\n\n")
  
  tax_info <- params$taxonomy_summary
  
  cat("- **Total Species Detected:** ", format_number(tax_info$total_species, 0), "\n")
  cat("- **Total Sequence Reads:** ", format_number(tax_info$total_reads, 0), "\n")
  cat("- **Average Reads per Sample:** ", format_number(tax_info$avg_reads_per_sample, 0), "\n")
  
  if(length(tax_info$top_species) > 0) {
    cat("- **Most Abundant Species:** ", paste(head(tax_info$top_species, 5), collapse = ", "), "\n")
  }
  
  cat("\n")
}
```

---

```{r alpha-section, results='asis', eval=("alpha" %in% params$selected_sections)}
cat("# Alpha Diversity Analysis\n\n")
cat("Alpha diversity measures the species diversity within individual samples. Higher alpha diversity typically indicates a more complex and potentially healthier microbial community.\n\n")

if(!is.null(params$alpha_stats)) {
  cat("## Summary Statistics by Group\n\n")
  
  # Create a comprehensive summary table
  alpha_table <- data.frame(
    Metric = rep(c("Shannon Index", "Simpson Index", "Observed Species"), 2),
    Group = c(rep("Patient", 3), rep("Control", 3)),
    Mean = c(
      params$alpha_stats$shannon_mean_patient,
      params$alpha_stats$simpson_mean_patient,
      params$alpha_stats$observed_mean_patient,
      params$alpha_stats$shannon_mean_control,
      params$alpha_stats$simpson_mean_control,
      params$alpha_stats$observed_mean_control
    ),
    `Standard Deviation` = c(
      params$alpha_stats$shannon_sd_patient,
      params$alpha_stats$simpson_sd_patient,
      params$alpha_stats$observed_sd_patient,
      params$alpha_stats$shannon_sd_control,
      params$alpha_stats$simpson_sd_control,
      params$alpha_stats$observed_sd_control
    ),
    check.names = FALSE
  )
  
  print(kable(alpha_table, 
              caption = "Alpha Diversity Summary Statistics by Group",
              digits = 3) %>%
        kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
        column_spec(2, bold = TRUE) %>%
        row_spec(which(alpha_table$Group == "Patient"), background = "#e3f2fd") %>%
        row_spec(which(alpha_table$Group == "Control"), background = "#e8f5e8"))
  
  cat("\n\n")
  
  # Statistical comparison
  cat("## Group Comparisons\n\n")
  
  shannon_diff <- params$alpha_stats$shannon_mean_patient - params$alpha_stats$shannon_mean_control
  simpson_diff <- params$alpha_stats$simpson_mean_patient - params$alpha_stats$simpson_mean_control
  observed_diff <- params$alpha_stats$observed_mean_patient - params$alpha_stats$observed_mean_control
  
  cat("**Key Findings:**\n\n")
  cat("- **Shannon Diversity:** ")
  if(shannon_diff > 0) {
    cat("Patients show higher diversity (difference: +", round(shannon_diff, 3), ")\n")
  } else {
    cat("Controls show higher diversity (difference: ", round(shannon_diff, 3), ")\n")
  }
  
  cat("- **Simpson Diversity:** ")
  if(simpson_diff > 0) {
    cat("Patients show higher diversity (difference: +", round(simpson_diff, 3), ")\n")
  } else {
    cat("Controls show higher diversity (difference: ", round(simpson_diff, 3), ")\n")
  }
  
  cat("- **Species Richness:** ")
  if(observed_diff > 0) {
    cat("Patients have more species on average (difference: +", round(observed_diff, 1), " species)\n")
  } else {
    cat("Controls have more species on average (difference: ", round(observed_diff, 1), " species)\n")
  }
  
  cat("\n\n")
  cat("**Interpretation:**\n")
  cat("- **Shannon Index:** Measures both richness and evenness. Range: 0-∞ (higher = more diverse)\n")
  cat("- **Simpson Index:** Probability that two randomly selected individuals belong to different species. Range: 0-1 (higher = more diverse)\n")
  cat("- **Observed Species:** Total number of different species detected in each sample\n\n")
}
```

```{r alpha-plot, eval=("alpha" %in% params$selected_sections), fig.width=10, fig.height=6}
# Check for alpha diversity plot
if (!is.null(params$plot_files) && !is.null(params$plot_files$alpha_plot)) {
  if (file_exists_and_valid(params$plot_files$alpha_plot)) {
    cat("## Alpha Diversity Visualization\n\n")
    knitr::include_graphics(params$plot_files$alpha_plot)
  } else {
    cat("## Alpha Diversity Visualization\n\n")
    cat("*Alpha diversity plot could not be generated.*\n\n")
    if (file.exists(params$plot_files$alpha_plot)) {
      cat("Plot file exists but appears to be empty or corrupted.\n\n")
    } else {
      cat("Plot file was not created.\n\n")
    }
  }
} else {
  cat("## Alpha Diversity Visualization\n\n")
  cat("*Alpha diversity plot not available - plot_files$alpha_plot is NULL.*\n\n")
}
```

```{r beta-section, results='asis', eval=("beta" %in% params$selected_sections)}
cat("\n\n# Beta Diversity Analysis\n\n")
cat("Beta diversity analysis examines the differences in species composition between samples, revealing patterns of community structure and relationships between different sample groups.\n\n")

cat("## Community Structure Analysis\n\n")
cat("Two complementary approaches are used to visualize beta diversity patterns:\n\n")
cat("1. **Principal Coordinate Analysis (PCoA):** Reduces the dimensionality of community data while preserving distances between samples\n")
cat("2. **Hierarchical Clustering:** Groups samples based on similarity, showing relationships in a tree-like structure\n\n")
```

```{r beta-pcoa-plot, eval=("beta" %in% params$selected_sections), fig.width=10, fig.height=8}
if (!is.null(params$plot_files) && !is.null(params$plot_files$beta_plot)) {
  if (file_exists_and_valid(params$plot_files$beta_plot)) {
    cat("### Principal Coordinate Analysis (PCoA)\n\n")
    cat("PCoA ordination showing the first two principal coordinates. Samples that cluster together have similar microbial compositions. Ellipses represent 95% confidence intervals for each group.\n\n")
    knitr::include_graphics(params$plot_files$beta_plot)
  } else {
    cat("### Principal Coordinate Analysis (PCoA)\n\n")
    cat("*PCoA plot could not be generated.*\n\n")
  }
} else {
  cat("### Principal Coordinate Analysis (PCoA)\n\n")
  cat("*PCoA plot not available.*\n\n")
}
```

```{r beta-dendrogram-plot, eval=("beta" %in% params$selected_sections), fig.width=14, fig.height=10}
if (!is.null(params$plot_files) && !is.null(params$plot_files$beta_dendrogram)) {
  if (file_exists_and_valid(params$plot_files$beta_dendrogram)) {
    cat("\n\n### Hierarchical Clustering Dendrogram\n\n")
    cat("Hierarchical clustering dendrogram showing relationships between samples based on Bray-Curtis dissimilarity. ")
    cat("The tree structure reveals which samples are most similar to each other. ")
    cat("Colored rectangles indicate major clusters identified by the algorithm.\n\n")
    knitr::include_graphics(params$plot_files$beta_dendrogram)
  } else {
    cat("\n\n### Hierarchical Clustering Dendrogram\n\n")
    cat("*Clustering dendrogram could not be generated.*\n\n")
  }
} else {
  cat("\n\n### Hierarchical Clustering Dendrogram\n\n")
  cat("*Clustering dendrogram not available.*\n\n")
}
```

```{r beta-interpretation, results='asis', eval=("beta" %in% params$selected_sections)}
cat("\n\n## Interpretation Guide\n\n")

cat("**Understanding the Plots:**\n\n")

cat("**PCoA Plot:**\n")
cat("- Points represent individual samples\n")
cat("- Distance between points reflects dissimilarity in microbial composition\n")
cat("- Closer points = more similar microbiomes\n")
cat("- Variance explained by each axis is shown in parentheses\n")
cat("- Group ellipses help visualize separation between sample types\n\n")

cat("**Clustering Dendrogram:**\n")
cat("- Height of branches indicates degree of dissimilarity\n")
cat("- Samples joined at lower heights are more similar\n")
cat("- Major clusters are highlighted with colored rectangles\n")
cat("- Ward.D2 linkage method minimizes within-cluster variance\n\n")

cat("**Biological Significance:**\n")
cat("- Clear separation between groups suggests distinct microbiome signatures\n")
cat("- Mixed clustering may indicate heterogeneity within groups\n")
cat("- Outlier samples may represent unique microbiome states or technical artifacts\n\n")
```


```{r taxonomy-section, results='asis', eval=("taxonomy" %in% params$selected_sections)}
cat("\n\n# Taxonomic Composition Analysis\n\n")
cat("Taxonomic composition analysis reveals the relative abundance of different microbial species across samples, providing insights into the dominant members of the microbial community and differences between sample groups.\n\n")

if(!is.null(params$taxonomy_summary)) {
  tax_info <- params$taxonomy_summary
  
  cat("## Key Findings\n\n")
  cat("- A total of **", format_number(tax_info$total_species, 0), " species** were identified across all samples\n")
  cat("- The analysis processed **", format_number(tax_info$total_reads, 0), " sequence reads**\n")
  cat("- Average sequencing depth was **", format_number(tax_info$avg_reads_per_sample, 0), " reads per sample**\n\n")
  
  if(length(tax_info$top_species) > 0) {
    cat("## Most Abundant Species\n\n")
    top_species_df <- data.frame(
      Rank = 1:length(tax_info$top_species),
      Species = tax_info$top_species
    )
    
    print(kable(top_species_df, 
                caption = "Top 10 Most Frequently Detected Species") %>%
          kable_styling(bootstrap_options = c("striped", "hover", "condensed")))
    cat("\n\n")
  }
}
```

```{r taxonomy-composition-plot, eval=("taxonomy" %in% params$selected_sections), fig.width=16, fig.height=10}
if (!is.null(params$plot_files) && !is.null(params$plot_files$taxonomy_plot)) {
  if (file_exists_and_valid(params$plot_files$taxonomy_plot)) {
    cat("## Species Composition Overview\n\n")
    cat("The stacked bar chart below shows the relative abundance of the most dominant species in each sample, grouped by sample type. Each color represents a different species, and the height of each segment represents its relative abundance within that sample.\n\n")
    knitr::include_graphics(params$plot_files$taxonomy_plot)
  } else {
    cat("## Species Composition Overview\n\n")
    cat("*Taxonomic composition plot could not be generated.*\n\n")
  }
} else {
  cat("## Species Composition Overview\n\n")
  cat("*Taxonomic composition plot not available.*\n\n")
}
```

```{r taxonomy-heatmap-plot, eval=("taxonomy" %in% params$selected_sections), fig.width=14, fig.height=10}
if (!is.null(params$plot_files) && !is.null(params$plot_files$taxonomy_heatmap)) {
  if (file_exists_and_valid(params$plot_files$taxonomy_heatmap)) {
    cat("\n\n## Taxonomic Abundance Heatmap\n\n")
    cat("The heatmap provides a comprehensive view of species abundance patterns across all samples. ")
    cat("Samples are clustered by similarity (columns) and species are grouped by abundance patterns (rows). ")
    cat("Color intensity represents log-transformed relative abundance, with darker colors indicating higher abundance.\n\n")
    
    cat("**Key Features:**\n")
    cat("- **Hierarchical clustering** groups similar samples and species together\n")
    cat("- **Sample annotations** (top bar) distinguish between patient and control groups\n")
    cat("- **Color scale** represents log10-transformed relative abundance values\n")
    cat("- **Dendrograms** show relationships between samples (top) and species (left)\n\n")
    
    knitr::include_graphics(params$plot_files$taxonomy_heatmap)
  } else {
    cat("\n\n## Taxonomic Abundance Heatmap\n\n")
    cat("*Taxonomic heatmap could not be generated.*\n\n")
  }
} else {
  cat("\n\n## Taxonomic Abundance Heatmap\n\n")
  cat("*Taxonomic heatmap not available.*\n\n")
}
```

```{r taxonomy-interpretation, results='asis', eval=("taxonomy" %in% params$selected_sections)}
cat("\n\n## Interpretation Guide\n\n")

cat("**Understanding the Visualizations:**\n\n")

cat("**Stacked Bar Chart:**\n")
cat("- Each bar represents one sample\n")
cat("- Colors represent different species\n")
cat("- Height of each colored segment = relative abundance of that species\n")
cat("- Samples are grouped by type (Patient vs Control) for easy comparison\n")
cat("- Only the most abundant species are shown individually; rare species are grouped as \"Other\"\n\n")

cat("**Abundance Heatmap:**\n")
cat("- Rows = species, Columns = samples\n")
cat("- Color intensity = species abundance (log-transformed)\n")
cat("- Clustering reveals patterns in microbial community structure\n")
cat("- Sample groupings (top annotation) help identify group-specific patterns\n")
cat("- Similar samples cluster together (shown by column dendrogram)\n")
cat("- Co-occurring species cluster together (shown by row dendrogram)\n\n")

cat("**Biological Insights:**\n")
cat("- **Dominant species** appear as consistent high-abundance bands across samples\n")
cat("- **Group-specific patterns** are visible when one sample type consistently shows different abundance patterns\n")
cat("- **Individual variation** is revealed by differences in abundance patterns between samples of the same type\n")
cat("- **Core microbiome** species appear in most samples, while **variable species** are present in only some samples\n\n")

if(!is.null(params$taxonomy_summary)) {
  tax_info <- params$taxonomy_summary
  
  cat("**Dataset-Specific Observations:**\n")
  cat("- This dataset contains", format_number(tax_info$total_species, 0), "distinct species\n")
  cat("- Average sequencing depth of", format_number(tax_info$avg_reads_per_sample, 0), "reads per sample provides")
  if(tax_info$avg_reads_per_sample > 10000) {
    cat(" excellent resolution for abundance analysis\n")
  } else if(tax_info$avg_reads_per_sample > 1000) {
    cat(" adequate resolution for abundance analysis\n")
  } else {
    cat(" limited resolution that may affect detection of rare species\n")
  }
  
  if(length(tax_info$top_species) > 0) {
    cat("- The most frequently detected species include:", paste(head(tax_info$top_species, 3), collapse = ", "), "\n")
  }
}
```

```{r phylo-section, results='asis', eval=("phylo" %in% params$selected_sections)}
cat("\n\n# Phylogenetic Analysis\n\n")
cat("Phylogenetic analysis incorporates evolutionary relationships between species to provide deeper insights into community structure and evolutionary diversity.\n\n")

cat("## Phylogenetic Tree\n\n")
cat("The phylogenetic tree shows the evolutionary relationships among the detected microbial species. Closely related species appear as neighboring branches, while distantly related species are separated by longer branch lengths.\n\n")
```

```{r phylo-plot, eval=("phylo" %in% params$selected_sections)}
if (!is.null(params$plot_files$phylo_plot) && file_exists_and_valid(params$plot_files$phylo_plot)) {
  knitr::include_graphics(params$plot_files$phylo_plot)
} else {
  cat("*Phylogenetic tree not available*\n")
}
```

```{r metadata-section, results='asis', eval=("metadata" %in% params$selected_sections)}
cat("\n\n# Metadata Analysis\n\n")
cat("Analysis of sample metadata provides context for understanding microbiome patterns in relation to host characteristics and experimental conditions. This section explores relationships between clinical variables and microbial community structure.\n\n")

if(!is.null(params$sample_summary)) {
  sample_info <- params$sample_summary
  
  cat("## Sample Demographics\n\n")
  
  # Create demographics table
  demo_data <- list()
  if(!identical(sample_info$age_range, "N/A")) {
    demo_data[["Age Range"]] <- sample_info$age_range
  }
  if(!identical(sample_info$gender_dist, "N/A")) {
    gender_summary <- paste(names(sample_info$gender_dist), 
                           sample_info$gender_dist, 
                           sep = ": ", collapse = "; ")
    demo_data[["Gender Distribution"]] <- gender_summary
  }
  demo_data[["Collection Period"]] <- sample_info$collection_period
  demo_data[["Total Samples"]] <- sample_info$total_samples
  demo_data[["Control Samples"]] <- sample_info$total_controls
  
  if(length(demo_data) > 0) {
    demo_df <- data.frame(
      Characteristic = names(demo_data),
      Value = unlist(demo_data),
      row.names = NULL
    )
    
    print(kable(demo_df, 
                caption = "Sample Characteristics Summary") %>%
          kable_styling(bootstrap_options = c("striped", "hover", "condensed")))
    cat("\n\n")
  }
}
```

```{r metadata-demographics-plot, eval=("metadata" %in% params$selected_sections), fig.width=10, fig.height=6}
if (!is.null(params$plot_files) && !is.null(params$plot_files$metadata_plot)) {
  if (file_exists_and_valid(params$plot_files$metadata_plot)) {
    cat("## Demographic Analysis\n\n")
    cat("Age and gender distribution analysis helps identify potential confounding factors and ensures balanced representation across sample groups.\n\n")
    knitr::include_graphics(params$plot_files$metadata_plot)
  }
}
```

```{r parallel-coordinates-plot, eval=("metadata" %in% params$selected_sections), fig.width=14, fig.height=8}
if (!is.null(params$plot_files) && !is.null(params$plot_files$parallel_plot)) {
  if (file_exists_and_valid(params$plot_files$parallel_plot)) {
    cat("\n\n## Multidimensional Metadata Analysis\n\n")
    cat("The parallel coordinates plot provides a comprehensive view of relationships between multiple metadata variables and microbial community characteristics. Each line represents one sample, and the path of the line across variables reveals patterns and correlations.\n\n")
    
    cat("**How to Read the Plot:**\n")
    cat("- Each vertical axis represents a different variable (scaled 0-1)\n")
    cat("- Each colored line represents one sample traced across all variables\n")
    cat("- Line color indicates sample group (Patient vs Control)\n")
    cat("- Parallel lines suggest correlation between variables\n")
    cat("- Crossing lines indicate different patterns between groups\n\n")
    
    knitr::include_graphics(params$plot_files$parallel_plot)
  } else {
    cat("\n\n## Multidimensional Metadata Analysis\n\n")
    cat("*Parallel coordinates plot could not be generated.*\n\n")
  }
} else {
  cat("\n\n## Multidimensional Metadata Analysis\n\n")
  cat("*Parallel coordinates plot not available.*\n\n")
}
```

```{r metadata-interpretation, results='asis', eval=("metadata" %in% params$selected_sections)}
cat("\n\n## Clinical and Lifestyle Correlations\n\n")

cat("**Understanding Metadata Relationships:**\n\n")

cat("**Demographic Patterns:**\n")
cat("- Age and gender distributions help identify potential biological confounders\n")
cat("- Balanced representation across groups strengthens analytical conclusions\n")
cat("- Demographic outliers may indicate samples requiring special consideration\n\n")

cat("**Parallel Coordinates Insights:**\n")
cat("- **Group Separation:** Clear separation of colored lines suggests distinct profiles between patient and control groups\n")
cat("- **Variable Correlations:** Variables where lines tend to be parallel show positive correlation\n")
cat("- **Outlier Detection:** Individual lines that deviate significantly from group patterns may represent outlier samples\n")
cat("- **Pattern Recognition:** Similar line shapes within groups suggest common phenotypic characteristics\n\n")

cat("**Clinical Significance:**\n")
cat("- Lifestyle factors (exercise, diet, alcohol) may influence microbiome composition\n")
cat("- Medical conditions and medications can have direct effects on microbial communities\n")
cat("- BMI and age are known correlates of microbiome diversity and composition\n")
cat("- Antibiotic usage history is a major factor affecting microbial community structure\n\n")

if(!is.null(params$sample_summary)) {
  sample_info <- params$sample_summary
  
  cat("**Dataset-Specific Observations:**\n")
  
  if(!identical(sample_info$age_range, "N/A")) {
    cat("- Age range of", sample_info$age_range, "years provides good representation across life stages\n")
  }
  
  if(sample_info$total_controls > 0) {
    control_percentage <- round((sample_info$total_controls / sample_info$total_samples) * 100, 1)
    cat("- Control samples represent", control_percentage, "% of the dataset, enabling robust comparative analysis\n")
  }
  
  cat("- Collection period spanning", sample_info$collection_period, "helps control for temporal effects\n")
}

cat("\n**Recommendations for Further Analysis:**\n")
cat("- Consider stratifying microbiome analyses by key demographic variables\n")
cat("- Include significant metadata variables as covariates in statistical models\n")
cat("- Investigate outlier samples for potential technical or biological explanations\n")
cat("- Validate group differences while controlling for confounding variables\n\n")
```

# Statistical Methods

## Data Processing

- **Quality Control:** Raw sequencing data underwent quality filtering and preprocessing
- **Taxonomic Assignment:** Species identification using reference databases
- **Rarefaction:** Samples were rarefied to normalize sequencing depth differences
- **Diversity Metrics:** Alpha and beta diversity calculated using established ecological indices

## Statistical Analysis

- **Alpha Diversity:** Shannon and Simpson indices calculated using vegan package in R
- **Beta Diversity:** Bray-Curtis dissimilarity with Principal Coordinate Analysis (PCoA)
- **Statistical Testing:** Appropriate parametric and non-parametric tests applied based on data distribution
- **Multiple Testing Correction:** False Discovery Rate (FDR) correction applied where appropriate

## Software and Packages

This analysis was performed using R statistical software with the following key packages:

- **vegan:** Community ecology analysis
- **ape:** Phylogenetic analysis
- **ggplot2:** Data visualization
- **dplyr:** Data manipulation
- **Shiny:** Interactive web application framework

---

# Conclusions and Recommendations

```{r conclusions, results='asis'}
cat("## Key Findings\n\n")

if(!is.null(params$sample_summary) && !is.null(params$taxonomy_summary)) {
  sample_info <- params$sample_summary
  tax_info <- params$taxonomy_summary
  
  cat("1. **Sample Diversity:** This dataset contains", sample_info$total_samples, "samples")
  if(sample_info$total_controls > 0) {
    cat(" including", sample_info$total_controls, "control samples")
  }
  cat("\n\n")
  
  cat("2. **Taxonomic Richness:** A total of", format_number(tax_info$total_species, 0), 
      "species were identified, indicating")
  if(tax_info$total_species > 100) {
    cat(" high taxonomic diversity")
  } else if(tax_info$total_species > 50) {
    cat(" moderate taxonomic diversity")  
  } else {
    cat(" limited taxonomic diversity")
  }
  cat("\n\n")
  
  cat("3. **Sequencing Depth:** Average sequencing depth of", 
      format_number(tax_info$avg_reads_per_sample, 0), "reads per sample provides")
  if(tax_info$avg_reads_per_sample > 10000) {
    cat(" excellent coverage for diversity analysis")
  } else if(tax_info$avg_reads_per_sample > 1000) {
    cat(" adequate coverage for diversity analysis")
  } else {
    cat(" limited coverage that may affect analysis sensitivity")
  }
  cat("\n\n")
}

cat("## Recommendations for Future Analysis\n\n")
cat("- **Statistical Power:** Consider additional samples if comparing between groups\n")
cat("- **Functional Analysis:** Investigate metabolic pathways and functional potential\n")
cat("- **Longitudinal Studies:** Track microbiome changes over time if applicable\n")
cat("- **Environmental Factors:** Include additional metadata for comprehensive analysis\n\n")
```

---

**Report Generated Using:** AGB Microbiome Analysis Platform  
**Generation Time:** `r format(Sys.time(), "%B %d, %Y at %I:%M %p")`  
**Analysis Version:** 1.0

*For questions regarding this analysis, please contact the bioinformatics team.*
