
---
title: "`r params$report_title`"
author: "AGB Microbiome Analysis Platform"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
  pdf_document:
    toc: true
    number_sections: true
    latex_engine: xelatex
  word_document:
    toc: true
    reference_docx: NULL
params:
  report_title: "Microbiome Analysis Report"
  report_comments: ""
  selected_sections: ["summary", "alpha", "beta"]
  sample_summary: NULL
  taxonomy_summary: NULL
  alpha_stats: NULL
  plot_files: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE, 
  warning = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 300,
  out.width = "100%"
)

# Load required libraries
suppressPackageStartupMessages({
  library(ggplot2)
  library(knitr)
  library(dplyr)
  library(kableExtra)
})

# Helper function to check if file exists and is not empty
file_exists_and_valid <- function(filepath) {
  if (is.null(filepath) || filepath == "" || is.na(filepath)) {
    return(FALSE)
  }
  file.exists(filepath) && file.info(filepath)$size > 0
}

# Helper function to format numbers
format_number <- function(x, digits = 2) {
  if(is.numeric(x)) {
    formatC(x, format = "f", digits = digits, big.mark = ",")
  } else {
    as.character(x)
  }
}
```

---

**Report Generated:** `r format(Sys.time(), "%B %d, %Y at %I:%M %p")`

```{r comments, results='asis'}
if (!is.null(params$report_comments) && params$report_comments != "") {
  cat("\n**Analysis Comments:** ", params$report_comments, "\n\n")
}
```

---

# Executive Summary

This comprehensive microbiome analysis report presents findings from high-throughput sequencing data analysis. The report includes diversity metrics, taxonomic composition, and statistical comparisons between sample groups.

```{r summary-stats, results='asis'}
if(!is.null(params$sample_summary)) {
  cat("## Dataset Overview\n\n")
  
  sample_info <- params$sample_summary
  
  cat("- **Total Samples Analyzed:** ", sample_info$total_samples, "\n")
  cat("- **Control Samples:** ", sample_info$total_controls, "\n")
  cat("- **Collection Period:** ", sample_info$collection_period, "\n")
  cat("- **Age Range:** ", sample_info$age_range, " years\n")
  
  if(!identical(sample_info$gender_dist, "N/A")) {
    cat("- **Gender Distribution:** ")
    gender_text <- paste(names(sample_info$gender_dist), 
                        sample_info$gender_dist, 
                        sep = ": ", collapse = ", ")
    cat(gender_text, "\n")
  }
  
  cat("\n")
}

if(!is.null(params$taxonomy_summary)) {
  cat("## Taxonomic Overview\n\n")
  
  tax_info <- params$taxonomy_summary
  
  cat("- **Total Species Detected:** ", format_number(tax_info$total_species, 0), "\n")
  cat("- **Total Sequence Reads:** ", format_number(tax_info$total_reads, 0), "\n")
  cat("- **Average Reads per Sample:** ", format_number(tax_info$avg_reads_per_sample, 0), "\n")
  
  if(length(tax_info$top_species) > 0) {
    cat("- **Most Abundant Species:** ", paste(head(tax_info$top_species, 5), collapse = ", "), "\n")
  }
  
  cat("\n")
}
```

---

```{r alpha-section, results='asis', eval=("alpha" %in% params$selected_sections)}
cat("# Alpha Diversity Analysis\n\n")
cat("Alpha diversity measures the species diversity within individual samples. Higher alpha diversity typically indicates a more complex and potentially healthier microbial community.\n\n")

if(!is.null(params$alpha_stats)) {
  cat("## Summary Statistics\n\n")
  
  alpha_stats <- params$alpha_stats
  
  # Create a summary table
  alpha_table <- data.frame(
    Metric = c("Shannon Index", "Simpson Index", "Observed Species"),
    Mean = c(alpha_stats$shannon_mean, alpha_stats$simpson_mean, alpha_stats$observed_mean),
    `Standard Deviation` = c(alpha_stats$shannon_sd, alpha_stats$simpson_sd, alpha_stats$observed_sd),
    check.names = FALSE
  )
  
  print(kable(alpha_table, 
              caption = "Alpha Diversity Summary Statistics",
              digits = 3) %>%
        kable_styling(bootstrap_options = c("striped", "hover", "condensed")))
  
  cat("\n\n")
  cat("**Interpretation:**\n")
  cat("- **Shannon Index:** Measures both richness and evenness. Range: 0-∞ (higher = more diverse)\n")
  cat("- **Simpson Index:** Probability that two randomly selected individuals belong to different species. Range: 0-1 (higher = more diverse)\n")
  cat("- **Observed Species:** Total number of different species detected in each sample\n\n")
}
```

```{r alpha-plot, eval=("alpha" %in% params$selected_sections)}
if (!is.null(params$plot_files$alpha_plot) && file_exists_and_valid(params$plot_files$alpha_plot)) {
  knitr::include_graphics(params$plot_files$alpha_plot)
} else {
  cat("*Alpha diversity plot not available*\n")
}
```

```{r beta-section, results='asis', eval=("beta" %in% params$selected_sections)}
cat("\n\n# Beta Diversity Analysis\n\n")
cat("Beta diversity analysis examines the differences in species composition between samples. This analysis helps identify patterns of community structure and relationships between different sample groups.\n\n")

cat("## Principal Coordinate Analysis (PCoA)\n\n")
cat("PCoA is an ordination method that reduces the dimensionality of community data while preserving the distances between samples. Samples that cluster together have similar microbial compositions.\n\n")
```

```{r beta-plot, eval=("beta" %in% params$selected_sections)}
if (!is.null(params$plot_files$beta_plot) && file_exists_and_valid(params$plot_files$beta_plot)) {
  knitr::include_graphics(params$plot_files$beta_plot)
} else {
  cat("*Beta diversity plot not available*\n")
}
```

```{r taxonomy-section, results='asis', eval=("taxonomy" %in% params$selected_sections)}
cat("\n\n# Taxonomic Composition Analysis\n\n")
cat("Taxonomic composition analysis reveals the relative abundance of different microbial species across samples, providing insights into the dominant members of the microbial community.\n\n")

if(!is.null(params$taxonomy_summary)) {
  tax_info <- params$taxonomy_summary
  
  cat("## Key Findings\n\n")
  cat("- A total of **", format_number(tax_info$total_species, 0), " species** were identified across all samples\n")
  cat("- The analysis processed **", format_number(tax_info$total_reads, 0), " sequence reads**\n")
  cat("- Average sequencing depth was **", format_number(tax_info$avg_reads_per_sample, 0), " reads per sample**\n\n")
  
  if(length(tax_info$top_species) > 0) {
    cat("## Most Abundant Species\n\n")
    top_species_df <- data.frame(
      Rank = 1:length(tax_info$top_species),
      Species = tax_info$top_species
    )
    
    print(kable(top_species_df, 
                caption = "Top 10 Most Frequently Detected Species") %>%
          kable_styling(bootstrap_options = c("striped", "hover", "condensed")))
    cat("\n\n")
  }
}
```

```{r taxonomy-plot, eval=("taxonomy" %in% params$selected_sections)}
if (!is.null(params$plot_files$taxonomy_plot) && file_exists_and_valid(params$plot_files$taxonomy_plot)) {
  knitr::include_graphics(params$plot_files$taxonomy_plot)
} else {
  cat("*Taxonomic composition plot not available*\n")
}
```

```{r phylo-section, results='asis', eval=("phylo" %in% params$selected_sections)}
cat("\n\n# Phylogenetic Analysis\n\n")
cat("Phylogenetic analysis incorporates evolutionary relationships between species to provide deeper insights into community structure and evolutionary diversity.\n\n")

cat("## Phylogenetic Tree\n\n")
cat("The phylogenetic tree shows the evolutionary relationships among the detected microbial species. Closely related species appear as neighboring branches, while distantly related species are separated by longer branch lengths.\n\n")
```

```{r phylo-plot, eval=("phylo" %in% params$selected_sections)}
if (!is.null(params$plot_files$phylo_plot) && file_exists_and_valid(params$plot_files$phylo_plot)) {
  knitr::include_graphics(params$plot_files$phylo_plot)
} else {
  cat("*Phylogenetic tree not available*\n")
}
```

```{r metadata-section, results='asis', eval=("metadata" %in% params$selected_sections)}
cat("\n\n# Metadata Analysis\n\n")
cat("Analysis of sample metadata provides context for understanding microbiome patterns in relation to host characteristics and experimental conditions.\n\n")

if(!is.null(params$sample_summary)) {
  sample_info <- params$sample_summary
  
  cat("## Sample Demographics\n\n")
  
  # Create demographics table
  demo_data <- list()
  if(!identical(sample_info$age_range, "N/A")) {
    demo_data[["Age Range"]] <- sample_info$age_range
  }
  if(!identical(sample_info$gender_dist, "N/A")) {
    gender_summary <- paste(names(sample_info$gender_dist), 
                           sample_info$gender_dist, 
                           sep = ": ", collapse = "; ")
    demo_data[["Gender Distribution"]] <- gender_summary
  }
  demo_data[["Collection Period"]] <- sample_info$collection_period
  demo_data[["Total Samples"]] <- sample_info$total_samples
  demo_data[["Control Samples"]] <- sample_info$total_controls
  
  if(length(demo_data) > 0) {
    demo_df <- data.frame(
      Characteristic = names(demo_data),
      Value = unlist(demo_data),
      row.names = NULL
    )
    
    print(kable(demo_df, 
                caption = "Sample Characteristics Summary") %>%
          kable_styling(bootstrap_options = c("striped", "hover", "condensed")))
    cat("\n\n")
  }
}
```

```{r metadata-plot, eval=("metadata" %in% params$selected_sections)}
if (!is.null(params$plot_files$metadata_plot) && file_exists_and_valid(params$plot_files$metadata_plot)) {
  knitr::include_graphics(params$plot_files$metadata_plot)
} else {
  cat("*Metadata analysis plot not available*\n")
}
```

# Statistical Methods

## Data Processing

- **Quality Control:** Raw sequencing data underwent quality filtering and preprocessing
- **Taxonomic Assignment:** Species identification using reference databases
- **Rarefaction:** Samples were rarefied to normalize sequencing depth differences
- **Diversity Metrics:** Alpha and beta diversity calculated using established ecological indices

## Statistical Analysis

- **Alpha Diversity:** Shannon and Simpson indices calculated using vegan package in R
- **Beta Diversity:** Bray-Curtis dissimilarity with Principal Coordinate Analysis (PCoA)
- **Statistical Testing:** Appropriate parametric and non-parametric tests applied based on data distribution
- **Multiple Testing Correction:** False Discovery Rate (FDR) correction applied where appropriate

## Software and Packages

This analysis was performed using R statistical software with the following key packages:

- **vegan:** Community ecology analysis
- **ape:** Phylogenetic analysis
- **ggplot2:** Data visualization
- **dplyr:** Data manipulation
- **Shiny:** Interactive web application framework

---

# Conclusions and Recommendations

```{r conclusions, results='asis'}
cat("## Key Findings\n\n")

if(!is.null(params$sample_summary) && !is.null(params$taxonomy_summary)) {
  sample_info <- params$sample_summary
  tax_info <- params$taxonomy_summary
  
  cat("1. **Sample Diversity:** This dataset contains", sample_info$total_samples, "samples")
  if(sample_info$total_controls > 0) {
    cat(" including", sample_info$total_controls, "control samples")
  }
  cat("\n\n")
  
  cat("2. **Taxonomic Richness:** A total of", format_number(tax_info$total_species, 0), 
      "species were identified, indicating")
  if(tax_info$total_species > 100) {
    cat(" high taxonomic diversity")
  } else if(tax_info$total_species > 50) {
    cat(" moderate taxonomic diversity")  
  } else {
    cat(" limited taxonomic diversity")
  }
  cat("\n\n")
  
  cat("3. **Sequencing Depth:** Average sequencing depth of", 
      format_number(tax_info$avg_reads_per_sample, 0), "reads per sample provides")
  if(tax_info$avg_reads_per_sample > 10000) {
    cat(" excellent coverage for diversity analysis")
  } else if(tax_info$avg_reads_per_sample > 1000) {
    cat(" adequate coverage for diversity analysis")
  } else {
    cat(" limited coverage that may affect analysis sensitivity")
  }
  cat("\n\n")
}

cat("## Recommendations for Future Analysis\n\n")
cat("- **Statistical Power:** Consider additional samples if comparing between groups\n")
cat("- **Functional Analysis:** Investigate metabolic pathways and functional potential\n")
cat("- **Longitudinal Studies:** Track microbiome changes over time if applicable\n")
cat("- **Environmental Factors:** Include additional metadata for comprehensive analysis\n\n")
```

---

**Report Generated Using:** AGB Microbiome Analysis Platform  
**Generation Time:** `r format(Sys.time(), "%B %d, %Y at %I:%M %p")`  
**Analysis Version:** 1.0

*For questions regarding this analysis, please contact the bioinformatics team.*
